---
title: "Demonstrating rflexdid on simulated data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demonstrating rflexdid on simulated data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
To regenerate rflexdid_demo.md (the GitHub-rendered version) after editing:

    Rscript -e 'devtools::install(quiet=TRUE); setwd("vignettes"); knitr::knit("rflexdid_demo.Rmd", output = "rflexdid_demo.md"); txt <- readLines("rflexdid_demo.md"); delims <- which(txt == "---"); if (length(delims) >= 2) txt <- txt[seq(delims[2] + 1, length(txt))]; first <- which(nchar(trimws(txt)) > 0)[1]; if (!is.na(first) && first > 1) txt <- txt[seq(first, length(txt))]; writeLines(txt, "rflexdid_demo.md")'

Run from the package root. devtools::install() is required so that
system.file("extdata", ..., package = "rflexdid") resolves correctly.
Then commit rflexdid_demo.md and figure/ alongside this file.
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Demonstrating rflexdid on simulated data

## The simulated dataset

The script `simulate_data.R` builds a repeated cross section of 20,000
adults surveyed across 40 counties over 10 years (2010–2019). Each year
brings a fresh sample of 50 individuals per county; no individual is
followed across time. Counties enact a sugar-sweetened-beverage (SSB)
excise tax in different years (the staggered cohorts: 2013, 2015, 2017,
plus a never-treated cohort labeled `0`). The outcome, `ssb_oz`, is daily SSB
consumption in fluid ounces.

The data-generating process imposes strict parallel pre-trends and known
treatment effects: an immediate jump `alpha_g` (-1.5 oz/day for the 2013
cohort, -1.0 for 2015, -0.5 for 2017) plus a smooth dynamic decline
`-0.20 e - 0.02 e^2` in event time `e`. So early-cohort counties
enacted larger taxes, and consumption keeps falling with longer exposure.

```{r data}
library(rflexdid)
library(ggplot2)
library(fixest)
df <- read.csv(system.file("extdata", "example_data.csv", package = "rflexdid"))
head(df)
table(df$cohort)
```

## Estimation

`flexdid()` fits the pooled OLS `FLEX` regression that includes treatment‑cell indicators together with group and time fixed effects.
When `specification = "lagsonly"` all pre-treatment periods are aggregated to form a reference period; when `specification = "lagsandleads"` the method includes separate indicators for each pre-treatment period, allowing for dynamic effects in the pre-period as well. In the latter case, the period just before treatment (`e = -1`) is the omitted reference period.
Covariates on the right‑hand side are interacted with the treatment‑cell indicators so that treatment effects can vary across cells; additive controls can instead be specified via `xnotinteracted`.

Identifying assumptions: (i) there must be at least one period in
which all units are untreated (no always‑treated units); (ii) conditional
parallel trends: untreated potential outcomes are captured by the included
group and time effects (and covariates when used); (iii) for repeated
cross‑sections, any compositional changes over time must be unrelated to
treatment timing conditional on covariates. See Deb et al. (2025) for formal
conditions.

```{r fit}
fit <- flexdid(ssb_oz ~ 1,
               data = df,
               tx = "treated",
               group = "county",
               time = "year",
               specification = "lagsandleads",
               vcov = "cluster",
               cluster = "county")

fit
```

## Overall ATET

A single number that averages cell-level effects across all
post-treatment observations. The DGP-implied truth (the mean of `tau`
over treated rows, using the same formula as `simulate_data.R`):

```{r truth}
alpha_g <- c(`2013` = -1.5, `2015` = -1.0, `2017` = -0.5, `0` = 0)
e <- df$year - df$cohort
tau_true <- ifelse(df$treated == 1,
                   alpha_g[as.character(df$cohort)] - 0.20 * e - 0.02 * e^2,
                   0)
mean(tau_true[df$treated == 1])
```

The `flexdid` estimate should be close to the DGP truth:

```{r overall}
atet(fit, type = "overall")
```

## Comparison to TWFE

For reference, the canonical two-way fixed-effects (TWFE) estimator
regresses the outcome on a single dummy for treatment, with group and
time fixed effects:

```{r twfe}
twfe <- feols(ssb_oz ~ treated | county + year,
              data = df,
              cluster = ~ county)

twfe
```

The TWFE coefficient on `treated` is noticeably attenuated relative to
the flexdid overall ATET above. This is the canonical
Goodman-Bacon (2021) result: under staggered treatment timing with
dynamic effects, TWFE includes "forbidden comparisons"
in which already-treated units serve as controls for later-treated
cohorts. When effects intensify with exposure, those negative-weighted
2 × 2 contrasts pull the estimate toward zero. The `flexdid` design
sidesteps the problem by estimating each cohort × group × time cell
separately and then aggregating only legitimate comparisons.

## Including covariates

Covariates placed on the RHS are interacted with the treatment‑cell
indicators, permitting treatment effects to vary with those covariates
across cells. In this DGP age does not moderate treatment, so the overall ATET barely moves; the standard error tightens slightly:

```{r include_age}
fit_age <- flexdid(ssb_oz ~ age,
                   data = df,
                   tx = "treated",
                   group = "county",
                   time = "year",
                   specification = "lagsandleads",
                   vcov = "cluster",
                   cluster = "county")

atet(fit_age, type = "overall")
```

Controls supplied via `xnotinteracted` enter additively,
absorbing mean differences but not allowing treatment‑effect heterogeneity.
This can improve precision when covariates are strong predictors of the outcome but do not moderate treatment effects. It can also help make a conditional parallel trends assumption more plausible without inflating the parameter count by interacting covariates with treatment‑cell indicators.

(Note: covariates that are constant within a group are collinear with the group fixed effects and will be dropped.)

```{r xnotinteracted}
fit2 <- flexdid(ssb_oz ~ 1,
                data = df,
                tx = "treated",
                group = "county",
                time = "year",
                specification = "lagsandleads",
                xnotinteracted = ~ region,
                vcov = "cluster",
                cluster = "county")

atet(fit2, type = "overall")
```


The remaining sections continue to use the baseline (`fit`, no
covariates) for clarity.

## Event study (`byexposure`)

The most useful diagnostic. Pre-period estimates (event time `e < 0`)
should be statistically indistinguishable from zero; post-period
estimates should grow more negative with exposure time, tracing out the
dynamic profile of the tax's effect. The base period `e = -1` is the
omitted reference category in the event‑study specification and is
therefore set to zero by construction.

```{r byexposure}
ax <- atet(fit, type = "byexposure", test = "pretrends")
ax
plot(ax)
```

The pre-test of parallel trends evaluates H₀: all pre-treatment ATETs = 0.
Since the DGP imposes strict parallel pre-trends, we do not reject — consistent
with zero pre-treatment differences.

## By treated cohort

Each treated cohort gets one ATET, averaging that cohort's
post-treatment cell effects. The estimates should rank
2013 (most negative) > 2015 > 2017 (least negative), mirroring the tax
magnitudes baked into the DGP.

```{r bycohort}
acoh <- atet(fit, type = "bycohort")
acoh
plot(acoh)
```

## By calendar time

A separate ATET for each calendar year, averaging across whichever
cohorts were already post-treatment in that year. Effects start near
zero in 2010–2012 (no cohort treated yet), turn negative in 2013
(2013 cohort enters), and grow more negative as the 2015 and 2017
cohorts roll in.

```{r bycalendar}
acal <- atet(fit, type = "bycalendar")
acal
plot(acal)
```

## By treated group (county)

A separate ATET for each treated county. Useful for spotting outlier
counties or for inspecting the within-cohort dispersion.

```{r bygroup}
agrp <- atet(fit, type = "bygroup")
plot(agrp)
```

## Group × event time (`byget`)

The most disaggregated view: one ATET per (county, event time) pair.
Restricting `values = 0:3` keeps the table readable.

```{r byget}
aget <- atet(fit, type = "byget", values = 0:3)
head(as.data.frame(aget), 12)
```

## Subgroup ATET via `for_expr`

Restrict the population over which the ATET is averaged. This is the
analogue of Stata's `for(female == 1)` option. The subgroup estimate
should be close to the overall estimate, since the DGP doesn't make
treatment effects depend on sex.

```{r for_expr}
atet(fit, type = "byexposure", for_expr = ~ female == 1)
```

## Group-level aggregation weights

`aggregationweight = "grouplevel"` reweights cells so that each county
contributes equally to the ATET, regardless of cell size. In this
simulation every county has exactly 500 observations, so the result
matches the default `obslevel` weighting; with unbalanced cell sizes
the two would differ.

```{r aggregationweight}
atet(fit, type = "overall", aggregationweight = "grouplevel")
```

## Wald test of homogeneity

`test = "equal"` tests the null hypothesis that all event-time effects
are equal — i.e., that the dynamic profile is flat. The DGP makes
effects strictly increase in exposure, so the test should reject.

```{r equal}
atet(fit, type = "byexposure", test = "equal")
```
